### 2015-05-29 ISWAG_PRESENTATION/PRESENTATION_ISWAG.TEX
iswag presentation

\documentclass[a4, 14pts]{seminar}
\usepackage[dvips]{pstcol}
\usepackage{semcolor}
\usepackage{sem-page}
\usepackage{slidesec}
\input{seminar.bug}
\input{seminar.bg2}
\usepackage{semhelv}
\usepackage{subfigure}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{longtable,array,enumerate}

\usepackage{multicol}

\usepackage[intlimits,sumlimits,namelimits]{amsmath}
\usepackage{amssymb}
\usepackage[mathscr]{eucal}
\usepackage{mathrsfs}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}

\usepackage[dvips]{graphicx}
\usepackage{epsf}
\usepackage{epsfig}
\usepackage{fancyhdr}
\usepackage{layout}

\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}

\usepackage{url}
\usepackage{xspace}

\newtheorem{corollary}{Corollaire}[section]
\newtheorem{definition}[corollary]{D\'{e}finition}
\newtheorem{equations}[corollary]{Equations}
\newtheorem{example}[corollary]{Exemple}
\newtheorem{lemma}[corollary]{Lemme}
\newtheorem{proposition}[corollary]{Proposition}
\newtheorem{remark}[corollary]{Remarque}
\newtheorem{theorem}[corollary]{Th\'{e}or\`{e}me}
%%% The above 7 commands are used in the following way:
%%% The definition environment, for example, is created by
%%% \begin{definition}\label{xxx}. . .\end{definition}
\newcommand{\mylabel}[1]{\label{#1}
            \ifx\undefined\stillediting
            \else \fbox{$#1$}\fi }
\newcommand{\BE}{\begin{equation}}
\newcommand{\BEQ}[1]{\BE\mylabel{#1}}
\newcommand{\EEQ}{\end{equation}}
\newcommand{\rfb}[1]{\mbox{\rm
   (\ref{#1})}\ifx\undefined\stillediting\else:\fbox{$#1$}\fi}
\newcommand{\half}   {{\frac{1}{2}}}
\newenvironment{matr}[1]{\left[ \begin{array}{#1}}{\end{array}
                         \right]}
%
\newfont{\Blackboard}{msbm10 scaled 1200}
\newcommand{\bl}[1]{\mbox{\Blackboard #1}}
\newfont{\roma}{cmr10 scaled 1200}

\newcommand{\propp}{{\hskip -2.2mm{\bf .}\hskip 3mm}}

\newcommand{\ovra}{\overrightarrow}
% fin des commandes de Marius

\fancyhf{}
\renewcommand\headrulewidth{0.2mm}
\renewcommand\footrulewidth{0.2mm}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}

\newcommand{\ROT}{{\text{Rot}}}
\newcommand{\Rot}{\vec{\text{Rot}}}
\newcommand{\rot}{\vec{\text{rot}}}

\newcommand{\dsp}{\displaystyle}

\newcommand\eads{\textsf{}\xspace}
\newcommand\eadsccr{\textsf{}\xspace}

\fancyhead[L]{\theslideheading}
\fancyhead[R]{\textcolor{blue}{\eadsccr \thepage }}
\fancyfoot[R]{\textcolor{blue}{}}
\fancyfoot[L]{\textcolor{blue}{Deauville, 03 June 2015 }}
\renewcommand\headwidth{\textwidth}
\autoslidemarginstrue

\newslideframe{%
   myscshadow}[\SemcolorFrameOps
   \psset{shadowcolor=green}]{\psshadowbox{#1}}
\newcommand{\ud}{\mathrm{d}}

\newcommand{\ds}{\displaystyle}
\slideframe{myscshadow}
\setlength\slideframewidth{.5pt}
\setlength\footskip{1cm}
\setlength\headsep{1cm}

%\renewcommand\makeslideheading[1]{\par
%   \framebox{\theslidesection. #1}\par
%}
%\renewcommand\makeslidesubheading[1]{\par
%   \leavevmode\hspace*{1cm}%
%   \alph{slidesubsection}. #1\par
%}

\definecolor{Pink}{rgb}{1.,0.,0.2}
\definecolor{Mygreen}{rgb}{0.3,0.7,0.6}

\renewcommand\makeslideheading[1]{\par
   \fcolorbox{black}{Mygreen}{\color{yellow}\theslidesection. #1}\par
}
\renewcommand\makeslidesubheading[1]{\par
   \leavevmode\hspace*{1cm}%
   \textcolor{Mygreen}{\alph{slidesubsection}. #1}\par
}

\makeatletter
\renewcommand\slide@contents{\par
   \begin{center}
        \Large \bfseries\textcolor{blue}{Plan}
   \end{center}
  \def\l@slide##1##2##3{%
    \@undottedtocline{1}{1.5em}{2.3em}{\slidenumberline{##1.}{##2}}{}}%
  \def\l@subslide##1##2##3{%
    \@undottedtocline{1}{1.5em}{3.3em}{\slidenumberline{\@alph{##1})}{##2}}{}}%
  \@startlos
}
\makeatother

\slideframe{none}
%\XgrDefinitFormat{*}{line=0,mark=0,ll=1}
%\XgrDefinitFormat{pr}{}
%\XgrDefinitFormat{ks}{line=-1,mark=8,ll=0.4}
\begin{document}
\slidepagestyle{fancy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{slide}
\thispagestyle{plain}
\begin{center}
\textcolor{blue}{\huge Internal mesh optimization}\\
\textcolor{blue}{\huge Semantic linking and siloing}\\
\textcolor{blue}{\huge Big data}\\
\vspace{0.25 cm}
\textbf{DUPREY Stéfan}\\
\small{\textbf{Cdiscount}\\

}
%\eadsccr
\end{center}
\end{slide}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\small{
\begin{slide}
\slidecontents
\renewcommand\theslideheading{Plan}
\end{slide}
\begin{slide}
\slideheading{Introduction}
\textbf{\textcolor{blue}{\large Objective : give you an overview of the meshing optimization done at Cdiscount, the first french e-commerce web site }}
\small{
\newline $\bullet$ Heuristic based optimization algorithm to push specific products 
\newline $\bullet$ An e-commerce pitch
\newline $\bullet$ Semantic similarity and constraining for shrinking our universe
\newline $\bullet$ Big data implementation
}
\end{slide}
\begin{slide}
\slideheading{Some notations}
Let $N\in\mathbb{N}$ be the number of nodes in our mesh.
\newline
Let $\left(X_i\right)_{i \in \left\{1,...,N\right\}}$ be the vertices (URLs) of our oriented graph.
\newline
Let $\left(G_{ij}\right)  \in \left\{0,1\right\}^{N\times N}$ be the adjacency matrix of our oriented graph.
\newline
Let here define f, a given data per URL, which gives a potentiality metrics for our vertices.
\begin{equation}
\begin{array}{ccccc}
f & : & \left(X_i\right)_{i \in \left\{1,...,N\right\}} & \to & \mathbb{R}^{+} \\
& & x & \mapsto & f(x) \\
\end{array}
\end{equation}
\end{slide}
\begin{slide}
\slideheading{In-rank computation}
We restrain the universe to our site where we compute the standard page-rank.
\newline
Initialization :
\begin{equation}
\forall u \ PR\left(u\right)=\frac{1}{N}
\end{equation} 
Iterative computation :
\begin{equation}
PR\left(u\right)= \frac{\left(1-c\right)}{N} + c \times \sum_{v \rightarrow u}\frac{PR\left(v\right)}{card\left(\left\{v\rightarrow u\right\}\right)}
\end{equation}
\end{slide}
\begin{slide}
\slideheading{Exhaustive brute force doesn't work}
For a $N=10^6$ millions URLs web site, we have $2^{N^{2}}$ with a 2048 bits mantissa, 256 bits exponent 
\newline
\scriptsize{
$2^{{10^6}^2}=$
\\ \newline 9.5762442314927432848050594956989483747127095675192905698213128517073583274396016675898
\\ \newline 714705184143146468453752442806484690561169975498415015777492655947375270159476651418975
\\ \newline 300707658547568802353384879419803574730952480197774380552040662758127609571333683703207
\\ \newline 910070247048194459504686986124786492353387550318495241621572271925127288273993787778380
\\ \newline 450774809611395810191417363401889038757182279484019203870177413318113073911418463615759
\\ \newline 647977538478560166958988721048687854280187283661925937530017243461145905573802314471888
\\ \newline 491758757162677684017424597014433418179115289463552630751896559312213624470617453325056
\\ \newline 5836008e+301029995663
}
\end{slide}
\begin{slide}
\slideheading{Picturing our smallest store : jewelery }
\begin{center}
\epsfig{file=page_rank.eps,height=6.5cm,width=8.cm}
\end{center}
\end{slide}
\begin{slide}
\slideheading{Heuristic based algorithm}
We want here to optimize the adequation of our mesh $\left(X_i\right)$ to our potentiality vector $f$.
\begin{equation}
\max_{\left(G_{ij}\right)  \in \left\{0,1\right\}^{N\times N}}\left\{ \sum^{N}_{i=1} trafic\left(X_i\right)\times pageRank(X_i)\right\}
\end{equation} 
\end{slide}
\begin{slide}
\slideheading{Genetic algorithm}
\textbf{\textcolor{blue}{\large What a genetic algorithm }}
\small{
\newline $\bullet$ Genetic algorithm mimics evolutionary biology to find approximate solutions to optimization problems 
\newline $\bullet$ Start with an initial generation of candidate solutions that are tested against the objecive function (fitness of the individual)
\newline $\bullet$ Subsequent generations evolve from the first through selection, crossover and mutation
\newline $\bullet$ The individual that best minimizes the given objective is returned as the ideal solution
}
\newline \textbf{\textcolor{blue}{\large Why a genetic algorithm }}
\small{
\newline $\bullet$ Lots of local minima to avoid
\newline $\bullet$ Non continuous universe, constraints and objective
\newline $\bullet$ Problem with noise and non-smooth data
}
\end{slide}
\end{document}