### 2015-05-29 ISWAG_PRESENTATION/PRESENTATION_ISWAG.TEX
iswag pres

	  \documentclass[a4, 14pts]{seminar}
	  \usepackage[dvips]{pstcol}
	  \usepackage{semcolor}
	  \usepackage{sem-page}
	  \usepackage{slidesec}
	  \input{seminar.bug}
	  \input{seminar.bg2}
	  \usepackage{semhelv}
	  \usepackage{subfigure}
	  \usepackage[latin1]{inputenc}
	  \usepackage[T1]{fontenc}
	  \usepackage{times}
	  \usepackage{longtable,array,enumerate}

	  \usepackage{multicol}

	  \usepackage[intlimits,sumlimits,namelimits]{amsmath}
	  \usepackage{amssymb}
	  \usepackage[mathscr]{eucal}
	  \usepackage{mathrsfs}
	  \usepackage{amsthm}
	  \usepackage{amsfonts}
	  \usepackage{amsmath}

	  \usepackage[dvips]{graphicx}
	  \usepackage{epsf}
	  \usepackage{epsfig}
	  \usepackage{fancyhdr}
	  \usepackage{layout}

	  \usepackage[french]{babel}
	  \usepackage[T1]{fontenc}
	  \usepackage[latin1]{inputenc}

	  \usepackage{url}
	  \usepackage{xspace}

	  \newtheorem{corollary}{Corollaire}[section]
	  \newtheorem{definition}[corollary]{D\'{e}finition}
	  \newtheorem{equations}[corollary]{Equations}
	  \newtheorem{example}[corollary]{Exemple}
	  \newtheorem{lemma}[corollary]{Lemme}
	  \newtheorem{proposition}[corollary]{Proposition}
	  \newtheorem{remark}[corollary]{Remarque}
	  \newtheorem{theorem}[corollary]{Th\'{e}or\`{e}me}
	  %%% The above 7 commands are used in the following way:
	  %%% The definition environment, for example, is created by
	  %%% \begin{definition}\label{xxx}. . .\end{definition}
	  \newcommand{\mylabel}[1]{\label{#1}
		      \ifx\undefined\stillediting
		      \else \fbox{$#1$}\fi }
	  \newcommand{\BE}{\begin{equation}}
	  \newcommand{\BEQ}[1]{\BE\mylabel{#1}}
	  \newcommand{\EEQ}{\end{equation}}
	  \newcommand{\rfb}[1]{\mbox{\rm
	    (\ref{#1})}\ifx\undefined\stillediting\else:\fbox{$#1$}\fi}
	  \newcommand{\half}   {{\frac{1}{2}}}
	  \newenvironment{matr}[1]{\left[ \begin{array}{#1}}{\end{array}
				  \right]}
	  %
	  \newfont{\Blackboard}{msbm10 scaled 1200}
	  \newcommand{\bl}[1]{\mbox{\Blackboard #1}}
	  \newfont{\roma}{cmr10 scaled 1200}

	  \newcommand{\propp}{{\hskip -2.2mm{\bf .}\hskip 3mm}}

	  \newcommand{\ovra}{\overrightarrow}
	  % fin des commandes de Marius

	  \fancyhf{}
	  \renewcommand\headrulewidth{0.2mm}
	  \renewcommand\footrulewidth{0.2mm}

	  \newcommand{\R}{\mathbb{R}}
	  \newcommand{\C}{\mathbb{C}}
	  \newcommand{\N}{\mathbb{N}}
	  \newcommand{\Z}{\mathbb{Z}}

	  \newcommand{\ROT}{{\text{Rot}}}
	  \newcommand{\Rot}{\vec{\text{Rot}}}
	  \newcommand{\rot}{\vec{\text{rot}}}

	  \newcommand{\dsp}{\displaystyle}

	  \newcommand\eads{\textsf{}\xspace}
	  \newcommand\eadsccr{\textsf{}\xspace}

	  \fancyhead[L]{\theslideheading}
	  \fancyhead[R]{\textcolor{blue}{\eadsccr \thepage }}
	  \fancyfoot[R]{\textcolor{blue}{}}
	  \fancyfoot[L]{\textcolor{blue}{Deauville, 03 June 2015 }}
	  \renewcommand\headwidth{\textwidth}
	  \autoslidemarginstrue

	  \newslideframe{%
	    myscshadow}[\SemcolorFrameOps
	    \psset{shadowcolor=green}]{\psshadowbox{#1}}
	  \newcommand{\ud}{\mathrm{d}}

	  \newcommand{\ds}{\displaystyle}
	  \slideframe{myscshadow}
	  \setlength\slideframewidth{.5pt}
	  \setlength\footskip{1cm}
	  \setlength\headsep{1cm}

	  %\renewcommand\makeslideheading[1]{\par
	  %   \framebox{\theslidesection. #1}\par
	  %}
	  %\renewcommand\makeslidesubheading[1]{\par
	  %   \leavevmode\hspace*{1cm}%
	  %   \alph{slidesubsection}. #1\par
	  %}

	  \definecolor{Pink}{rgb}{1.,0.,0.2}
	  \definecolor{Mygreen}{rgb}{0.3,0.7,0.6}

	  \renewcommand\makeslideheading[1]{\par
	    \fcolorbox{black}{Mygreen}{\color{yellow}\theslidesection. #1}\par
	  }
	  \renewcommand\makeslidesubheading[1]{\par
	    \leavevmode\hspace*{1cm}%
	    \textcolor{Mygreen}{\alph{slidesubsection}. #1}\par
	  }

	  \makeatletter
	  \renewcommand\slide@contents{\par
	    \begin{center}
		  \Large \bfseries\textcolor{blue}{Plan}
	    \end{center}
	    \def\l@slide##1##2##3{%
	      \@undottedtocline{1}{1.5em}{2.3em}{\slidenumberline{##1.}{##2}}{}}%
	    \def\l@subslide##1##2##3{%
	      \@undottedtocline{1}{1.5em}{3.3em}{\slidenumberline{\@alph{##1})}{##2}}{}}%
	    \@startlos
	  }
	  \makeatother

	  \slideframe{none}
	  %\XgrDefinitFormat{*}{line=0,mark=0,ll=1}
	  %\XgrDefinitFormat{pr}{}
	  %\XgrDefinitFormat{ks}{line=-1,mark=8,ll=0.4}
	  \begin{document}
	  \slidepagestyle{fancy}
	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	  \begin{slide}
	  \thispagestyle{plain}
	  \begin{center}
	  \textcolor{blue}{\huge Internal mesh optimization}\\
	  \textcolor{blue}{\huge Semantic linking and siloing}\\
	  \textcolor{blue}{\huge Big data}\\
	  \vspace{0.25 cm}
	  \textbf{DUPREY Stéfan}\\
	  \small{\textbf{Cdiscount}\\

	  }
	  %\eadsccr
	  \end{center}
	  \end{slide}
	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	  {\small{
	  \begin{slide}
	  \slidecontents
	  \renewcommand\theslideheading{Plan}
	  \end{slide}
	  \begin{slide}
	  \slideheading{Introduction}
	  \textbf{\textcolor{blue}{\large Objective : give you an overview of the meshing optimization done at Cdiscount, the first french e-commerce web site }}
	  \small{
	  \newline $\bullet$ Heuristic based optimization algorithm to push specific products 
	  \newline $\bullet$ An e-commerce pitch
	  \newline $\bullet$ Semantic similarity and constraining for shrinking our universe
	  \newline $\bullet$ Big data implementation
	  }
	  \end{slide}
	  \begin{slide}
	  \slideheading{Heuristic}
	  \textbf{\textcolor{blue}{\large Some notations}}\newline
	  Let $N\in\mathbb{N}$ be the number of nodes in our mesh.
	  \newline
	  Let $\left(X_i\right)_{i \in \left\{1,...,N\right\}}$ be the vertices (URLs) of our oriented graph.
	  \newline
	  Let $\left(G_{ij}\right)  \in \left\{0,1\right\}^{N\times N}$ be the adjacency matrix of our oriented graph.
	  \newline
	  Let here define f, a given data per URL, which gives a potentiality metrics for our vertices.
	  \begin{equation}
	  \begin{array}{ccccc}
	  f & : & \left(X_i\right)_{i \in \left\{1,...,N\right\}} & \to & \mathbb{R}^{+} \\
	  & & x & \mapsto & f(x) \\
	  \end{array}
	  \end{equation}
	  \end{slide}
	  \begin{slide}
	  \textbf{\textcolor{blue}{\large In-rank}}\newline      
	  We restrain the universe to our site where we compute the standard page-rank.
	  \newline
	  Initialization :
	  \begin{equation}
	  \forall u \ PR\left(u\right)=\frac{1}{N}
	  \end{equation} 
	  Iterative computation :
	  \begin{equation}
	  PR\left(u\right)= \frac{\left(1-c\right)}{N} + c \times \sum_{v \rightarrow u}\frac{PR\left(v\right)}{card\left(\left\{v\rightarrow u\right\}\right)}
	  \end{equation}
	  \end{slide}
	  \begin{slide}
	  \textbf{\textcolor{blue}{\large Our heuristic}}\newline 
	  We want here to optimize the adequacy of our mesh $\left(X_i\right)$ to our potentiality vector $f$.
	  We here postulate the following heuristic to assess the relevance of a mesh :
	  \begin{equation}
	  \max_{\left(G_{ij}\right)  \in \left\{0,1\right\}^{N\times N}}\left\{ \sum^{N}_{i=1} trafic\left(X_i\right)\times pageRank(X_i)\right\}
	  \end{equation} 
	  \end{slide}

	  \begin{slide}
	  \slideheading{Metaheuristics}    
	  \textbf{\textcolor{blue}{\large Exhaustive brute force doesn't work}}\newline
	  For a $N=10^6$ millions URLs web site, we have $2^{N^{2}}$ with a 2048 bits mantissa, 256 bits exponent 
	  \newline
	  \scriptsize{
	  $2^{{10^6}^2}=$
	  \\ \newline 9.5762442314927432848050594956989483747127095675192905698213128517073583274396016675898
	  \\ \newline 714705184143146468453752442806484690561169975498415015777492655947375270159476651418975
	  \\ \newline 300707658547568802353384879419803574730952480197774380552040662758127609571333683703207
	  \\ \newline 910070247048194459504686986124786492353387550318495241621572271925127288273993787778380
	  \\ \newline 450774809611395810191417363401889038757182279484019203870177413318113073911418463615759
	  \\ \newline 647977538478560166958988721048687854280187283661925937530017243461145905573802314471888
	  \\ \newline 491758757162677684017424597014433418179115289463552630751896559312213624470617453325056
	  \\ \newline 5836008e+301029995663
	  }
	  \end{slide}
	  \begin{slide}
	  \textbf{\textcolor{blue}{\large Picturing our smallest store : jewelery}}\newline  
	  \begin{center}
	  \epsfig{file=page_rank.eps,height=6.5cm,width=8.cm}
	  \end{center}
	  \end{slide}
	  \begin{slide}
	  \textbf{\textcolor{blue}{\large Metaheuristics}}\newline
	  We here have to maximize over all possible graphs. 
	  We have to find a clever global optimization methodology : metaheuristics such as global search, multistart, particle swarm, simulated annealing or genetic algorithm are all good candidates.
	  \newline 
	  \textbf{\textcolor{black}{\large What a genetic algorithm}}
	  \scriptsize{
	  \newline $\bullet$ Genetic algorithm mimics evolutionary biology to find approximate solutions to optimization problems 
	  \newline $\bullet$ Start with an initial generation of candidate solutions that are tested against the objecive function (fitness of the individual)
	  \newline $\bullet$ Subsequent generations evolve from the first through selection, crossover and mutation
	  \newline $\bullet$ The individual that best minimizes the given objective is returned as the ideal solution
	  }
	  \newline \textbf{\textcolor{black}{\large Why a genetic algorithm}}
	  \scriptsize{
	  \newline $\bullet$ Lots of local minima to avoid
	  \newline $\bullet$ Non continuous universe, constraints and objective
	  \newline $\bullet$ Problem with noise and non-smooth data
	  }
	  \end{slide}
	  \begin{slide}
	  \textbf{\textcolor{blue}{\large Cleverly evolving through our universe}}\newline
	    \begin{center}
	    \textbf{\large Child spawning from 2 parents crossover}
	    \epsfig{file=crossover.eps,height=2cm,width=4cm}
	    \end{center}
	    \begin{center}
	    \textbf{\large Mutation of an individual}
	    \end{center}
	    \begin{center}
	    \epsfig{file=mutation.eps,height=0.4cm,width=4cm}
	    \end{center}
	  \end{slide}
	  \begin{slide}
	  \slideheading{Shrinking our universe with siloing and semantic similarity}
	    \textbf{\textcolor{blue}{\large Siloing and links categorizing}}\newline
	    \begin{center}
	    \epsfig{file=website_siloing.eps,height=4cm,width=6cm}
	    \end{center}
	  \begin{center}
	    \epsfig{file=links_categorization.eps,height=1cm,width=6cm}
	    \end{center}
	  \end{slide}
	  \begin{slide}
	    \textbf{\textcolor{blue}{\large Semantic similarity}}\newline
	  We slack our universe by allowing links only between semantically similar URLs :
	  \begin{equation}
	  \max_{\left(G_{ij}\right)  \in \left\{0,1\right\}^{N\times N}\ G_{ij}=0 \ if \ CS(ij)\leq t}\left\{ \sum^{N}_{i=1} f\left(X_i\right)\times PR(X_i)\right\}
	  \end{equation}
	  where $CS(ij)$ is a semantic distance between the two linked pages $i$ and $j$.
	  \newline $CS(ij)$ can be defined very easily as the scalar product of the tf/idf vectors of the
	  product descriptions whose weight are defined by the well known formula :
	  \begin{equation}
	  w_{ik}=\frac{tf_{ik}\log\left( \frac{N}{n_k}\right)}{\sqrt{\sum_{k=1}^t}(tf_{ik})^2 \log\left(\frac{N}{n_k}\right)^2}
	  \end{equation}
	  \end{slide}
	  \begin{slide}
	  \slideheading{Same problem with an e-commerce flavor}
	\begin{equation}
	\sum_{i\in Keywords} SV\left( i\right)\times CTR\left(position\left(i\right)\right)\times CR\left(i\right)\times P\left(i\right)
	\end{equation}
	where $position\left(i\right)$ is the estimated position in search engine results coming from the modification of our new mesh,
	\\\newline
	$SV\left(i\right)$ is the search volume for the keywords $i$ estimated by the search engine.
	\\\newline
	and $CTR\left(i\right)$ is the click through rate for an URL at the position place in the search engine results.
	\\\newline
	  \end{slide}
	  \begin{slide}
	  \slideheading{Prototyping example}
	    $$
    \left(X_i\right)=\{'home','metalist1','metalist2',
    $$
    $$
    'list1','list2','list3','list4',
    $$
    $$
    'product1','product2',...,'product32'\};
    $$
    $$
    f\left(X_i\right) = [100;80;55;$$
    $$
    40;35;25;20;
    $$
    $$
    4;5;...;3]
    $$
      \begin{center}
    \textbf{\large Initial population}
    \end{center}
    \begin{center}
    \epsfig{file=initial_population.eps,height=5cm,width=5cm}
    \end{center}
      \begin{center}
    \textbf{\large Population at convergence}
    \end{center}
    \begin{center}
    \epsfig{file=converged_population.eps,height=5cm,width=5cm}
    \end{center}
    \begin{center}
    \textbf{\large Best chromosome fitness function evolution}
    \end{center}
    \begin{center}
    \epsfig{file=best_element_evolution.eps,height=5cm,width=5cm}
    \end{center}
    The structure found is as we expected arborescent : home linking to metalists, metalists linking to lists and lists to products.
    \begin{center}
    \textbf{\large Optimal site structure}
    \end{center}
    \begin{center}
    \epsfig{file=site_structure.eps,height=5cm,width=5cm}
    \end{center}
	  \end{slide}
	  \begin{slide}
	    \slideheading{Big data implementation : Neo4j and Spark GraphX}
	    \begin{center}
	    \epsfig{file=neo4j_maze_runner.eps,height=3cm,width=8cm}
	    \end{center}
	  \end{slide}
	  \end{document}