### 2022-12-21 OrderBook/main.tex

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage[numbers]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}
\title{Risky order book detection}

\author{\IEEEauthorblockN{Stefan Duprey}
\IEEEauthorblockA{\textit{SwissBorg} \\
stefan@swissborg.com}
}

\maketitle

\begin{abstract}
In order to prevent the occurrence of high slippage orders when executing on behalf of our retail customer, we must detect risky order book and trade accordingly.
\begin{description}
  \item[$\bullet$]Detect inflexion points in the order book
  \item[$\bullet$]Cap maximum limit order price
  \item[$\bullet$]Expand time window to pass maker below the max price
  \item[$\bullet$]Fill with treasury the order part that would occur to big a slippage
\end{description}

\end{abstract}
\section{Snapshot only, no comparative historical data}
\subsection{Slippage definition}
\begin{center}
\begin{tabular}{||c c c||} 
 \hline
 Depth & Price & Quantity \\ [0.5ex] 
 \hline\hline
 ask depth n & \hdots & \hdots  \\ 
 \hline
 ... & &  \\
 \hline
 ask depth 2 & \hdots & \hdots  \\ 
 \hline
 best ask & \hdots & \hdots  \\ 
 \hline
 \hline
 best bid & \hdots & \hdots  \\ 
 \hline
 bid depth 2 & \hdots & \hdots  \\ 
 \hline
 ... & &  \\
 \hline
 bid depth n & \hdots & \hdots  \\ 
 \hline
 \end{tabular}
\end{center}
Let's define:
\begin{equation}
P_{mid\ market} = \frac{P_{best\ bid} + P_{best\ ask}}{2}    
\end{equation}
Let's define for an amount to pass $A$ the maximal depth $d$ at which the amount to pass would be filled if we ripped the order book.
\begin{equation}
d^* = \sup\left\{{d / \sum\limits_{j = 1}^{d} q_j < A}\right\}+1
\end{equation}
The execution price is then defined by the average price over all the depth up to $d^*$ (the last depth not being totally ripped off, just enough so that the $A$ amount is reached).\\
The execution price is then defined:
\begin{equation}
P_{execution}(A) =  \frac{\sum\limits_{j = 1}^{d^*(A)} q_j * p_j}{\sum\limits_{j=1}^{d^*(S)} q_j}
\end{equation}
And finally the slippage definition:
\begin{equation}
slippage(A) = \frac{P_{execution}(A) - P_{mid\ market}}{P_{mid\ market}}
\end{equation}
\subsection{Slippage acceleration}
We define the slippage acceleration as :
\begin{equation}
slippage\_acceleration(A)=\frac{d \left(slippage(A)\right)}{dA}  
\end{equation}

\subsection{Rules and absolute thresholds for slippage}

\susubbsection{Absolute threshold definition}
Let's denote S the slippage vector for different amounts:
$$
S =\begin{pmatrix} 
	S_{bid}(100) \\
	S_{ask}(100) \\
	S_{bid}(500) \\
	S_{ask}(500) \\
	\vdots\\
	S_{bid}(50000) \\
	S_{ask}(50000) \\
	S_{bid}(100000) \\
	S_{ask}(100000) \\
	\end{pmatrix}
$$
We define the absolute thresholds as the x\% quantile from the S vector distribution:
\begin{equation}
t =\begin{pmatrix} 
	t_{bid}(100) \\
	t_{ask}(100) \\
	t_{bid}(500) \\
	t_{ask}(500) \\
	\vdots\\
	t_{bid}(50000) \\
	t_{ask}(50000) \\
	t_{bid}(100000) \\
	t_{ask}(100000) \\
	\end{pmatrix}
\end{equation}
where $$\mathbb{P}\left[\lvert S(A) \rvert < t(A) \right] = x\% = 90\%$$ where $x$ is fixed (90\%)
\subsection{Common rule for all amounts}
We assess the riskiness of an order book with the following metrics:
\begin{equation}
R = \frac{\sum\limits_{way \in bid/ask}\sum\limits_{a_i} q_i * 1_{s_{way}(a_i)\ge t(a_i)}}{\sum\limits_{bid/ask,i} q_i}
\end{equation}
where $ (a_i)_{i \in [1,..,n]} = [100, 500, \hdots, 50000,100000]$ is the predefined range of amount 
and $q_i$ is a weighting here to emphasize that slower amount threshold break are more important than bigger amount. We propose first a linear weighting: if $a_i = [100,500,1000]$ then $q_i = [3,2,1]$
$$
q_i = range(len(a_i),1,-1)
$$
\subsection{Rule for one amount only}
We assess the riskiness of an order book for a single amount to pass $a$ with the following metrics:
\begin{equation}
R =  1_{s(a)\ge t(a)}
\end{equation}
where $a$ is the predefined amount. The order book is deemed risky if its slippage for the amount a is above the x\% threshold.
\subsection{Rule subsequent action}
If the order book slippage for the asked amount a is above the the absolute threshold $t(a)$:\\
Find $a^*$ such that $a^*<a$ and $slippage(a^*) = t(a)$ (possible as the slippage is an increasing function of the amount).\\
Fill the order till $a^$*, and then 2 solutions : 
\begin{description}
  \item[$\bullet$] Fill the other part with the treasury.
  \item[$\bullet$] Wait for the market maker to bring back liquidity below the last execution price
\end{description}

\section{Rolling thresholds}
We do the exact same analysis, but do calibrate the thresholds with quantile at 90\% from a rolling distribution.\\
The rolling window size must be defined and the quantiles recomputed dynamically.
\section{Dynamical historical thresholds}
Algorithmic is a tough field with really advanced and sophisticated players.\\
Snapshot algorithms are just a first simple step.\\
The next step would a dynamical study of the order book to assess slippage parameters due to order submissions.\\
Order book modelisation as VAR/VEC process can allow us to 
The dynamical part is the harder one. It implies to submit best bid/ask limit orders and inspect the repercussion of those depth quantities change over the order book.\\
This is actually where we found the most litterature about.

\section{Conclusion}

\bibliographystyle{IEEEtranN}
\bibliography{bib}
\nocite{*}
\end{document}

% to do
% Expliquer l'importance de la réduction de dimension dans ML
% Expliquer la diff CCA PCA pour ML